Seed ecology of European mesic meadows

### 2021-10-22 doc/manuscript.Rmd
ms accepted ☘️

---
title: "RESEARCH IN CONTEXT: Seed ecology of European mesic meadows"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: annals-of-botany.csl
link-citations: yes
always_allow_html: true
---

@@

Running title: Seed ecology of European mesic meadows

@@

# ABSTRACT (300 words)

@@

@@

@@

@@

@@

@@

The original data, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/meadows. Upon publication, a version of record of the repository will be deposited in Zenodo.

@@

# FIGURES

```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Diversity of dispersal units in plant species from mesic meadows: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadensis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (o) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
knitr::include_graphics("../results/figures/Fig1.png")
```

```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Morphology of mesic meadow seeds. The two panels on the left show the probability densities of species values for seed mass and seed number, log-transformed for ease of visualization. The three horizontal lines within the probability densities represent the first quartile, the median and the third quartile of the values. The panel on the right shows values of seed length and width obtained by image analysis, with each point being a seed. In all cases, data is divided between the grasses (Poaceae), legumes (Fabaceae) and the other plant families."}
knitr::include_graphics("../results/figures/Fig2.png")
```

```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Germination cues in mesic meadows. Effect of germination cues simulated in the laboratory over the final germination proportions of mesic meadow seeds. Dots indicate the posterior mean of the effect size for each cue, and whiskers the 95% credible interval of the effect size. The line of zero-effect is shown: when a credible interval overlaps with the zero-effect line, the effect can be regarded as non-significant. In separate panels, the figure shows the results of a general model including data for all species, plus specific models for the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families."}
knitr::include_graphics("../results/figures/Fig3.png")
```

```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Environment and seed traits are separate axis of variation in mesic meadows. Principal Component Analysis ordination of mesic meadow species considering their environmental preferences and their seed traits. Each point is a species, coloured by the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families. Labels indicate the contribution of the variables to the axes: grey-background labels for environmental preferences, and white-background labels for seed traits. Environmental preferences were calculated as species niche centroids (SNCs) for the minimal temperature of the coldest month (bio06), precipitation of the driest month (bio14), and soil pH. To calculate the SNCs, a vegetation dataset of mesic meadows of the Iberian Peninsula was used. The seed traits are seed mass and the germination relative indices for the response to scarification, stratification, average germination temperature (temperature), alternating temperature (alternating) and light. All environmental preferences aligned to the first axis, while seed traits aligned to the second axis, showing that environment and seed traits are separate axis of variation in mesic meadows. Stratification and average germination temperature showed very low variation, in accordance with their small effect on Figure 2."}
knitr::include_graphics("../results/figures/Fig4.png")
```

```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Higher germinability in mesic meadows compared to other herbaceous plant communities. Principal Component Analysis ordination of species responses to three seed germination treatments (fresh seeds germinated at 30/20, 22/12 and 14/4 ºC in a 12/12h light-darkness photoperiod). Species are grouped by their habitat, with points showing the centroid for each habitat group, and rays linking the centroid with the position of each species in the group. Labels indicate the contribution of the germination treatments to the axes. All species had been collected in herbaceous vegetation types of the Cantabrian Mountains of Spain and the neighbouring coast. All seeds were untreated, except for scarification, which was applied routinely to all botanical families presenting physical dormancy. Germination treatments consisted in 12/12 h periods with a warmer phase in light and a cooler phase in darkness. The position of mesic meadows compared to the other habitats indicates that their species tended to have higher germinability even if untreated (i.e. less seed dormancy) and were more neutral in their thermal requirements."}
knitr::include_graphics("../results/figures/Fig5.png")
```

>>>>

---
title: "Seed ecology of European mesic meadows"
output: github_document
bibliography: refs.bib
csl: annals-of-botany.csl
link-citations: yes
always_allow_html: true
---

@@

@@

# ABSTRACT

@@

```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Diversity of dispersal units in plant species from mesic meadows: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadensis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (o) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
knitr::include_graphics("../results/figures/Fig1.png")
```

@@

```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 2** Morphology of mesic meadow seeds. The two panels on the left show the probability densities of species values for seed mass and seed number, log-transformed for ease of visualization. The three horizontal lines within the probability densities represent the first quartile, the median and the third quartile of the values. The panel on the right shows values of seed length and width obtained by image analysis, with each point being a seed. In all cases, data is divided between the grasses (Poaceae), legumes (Fabaceae) and the other plant families."}
knitr::include_graphics("../results/figures/Fig2.png")
```

@@

```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 3** Germination cues in mesic meadows. Effect of germination cues simulated in the laboratory over the final germination proportions of mesic meadow seeds. Dots indicate the posterior mean of the effect size for each cue, and whiskers the 95% credible interval of the effect size. The line of zero-effect is shown: when a credible interval overlaps with the zero-effect line, the effect can be regarded as non-significant. In separate panels, the figure shows the results of a general model including data for all species, plus specific models for the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families."}
knitr::include_graphics("../results/figures/Fig3.png")
```

@@

```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 4** Environment and seed traits are separate axis of variation in mesic meadows. Principal Component Analysis ordination of mesic meadow species considering their environmental preferences and their seed traits. Each point is a species, coloured by the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families. Labels indicate the contribution of the variables to the axes: grey-background labels for environmental preferences, and white-background labels for seed traits. Environmental preferences were calculated as species niche centroids (SNCs) for the minimal temperature of the coldest month (bio06), precipitation of the driest month (bio14), and soil pH. To calculate the SNCs, a vegetation dataset of mesic meadows of the Iberian Peninsula was used. The seed traits are seed mass and the germination relative indices for the response to scarification, stratification, average germination temperature (temperature), alternating temperature (alternating) and light. All environmental preferences aligned to the first axis, while seed traits aligned to the second axis, showing that environment and seed traits are separate axis of variation in mesic meadows. Stratification and average germination temperature showed very low variation, in accordance with their small effect on Figure 2."}
knitr::include_graphics("../results/figures/Fig4.png")
```

@@

```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 5** Higher germinability in mesic meadows compared to other herbaceous plant communities. Principal Component Analysis ordination of species responses to three seed germination treatments (fresh seeds germinated at 30/20, 22/12 and 14/4 ºC in a 12/12h light-darkness photoperiod). Species are grouped by their habitat, with points showing the centroid for each habitat group, and rays linking the centroid with the position of each species in the group. Labels indicate the contribution of the germination treatments to the axes. All species had been collected in herbaceous vegetation types of the Cantabrian Mountains of Spain and the neighbouring coast. All seeds were untreated, except for scarification, which was applied routinely to all botanical families presenting physical dormancy. Germination treatments consisted in 12/12 h periods with a warmer phase in light and a cooler phase in darkness. The position of mesic meadows compared to the other habitats indicates that their species tended to have higher germinability even if untreated (i.e. less seed dormancy) and were more neutral in their thermal requirements."}
knitr::include_graphics("../results/figures/Fig5.png")
```

@@

The original data, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/meadows. A version of record of the repository is deposited in Zenodo.

@@
